### 2015-07-20 motivation.tex
Name change spack-paper-sc15 -> spack-sc15
Add ROCKS reference.
tweaks to motivation

\paragraph{Traditional package managers}
Package managers automate the installation of complex sets of software packages.
{\it Binary package managers} such as RPM, yum, and
APT~\cite{foster+:rpm03,silva:apt01,yum} are integrated with most
OS distributions, and they are used to ensure that dependencies
are installed before packages that require them.
These tools largely solve the problem of managing a {\it single} software
stack, which works well for the baseline OS and drivers, which are
common to all applications on a system.
These tools assume that each package only has a single version
and most of the tools install packages in a single, inflexible location.
To install multiple configurations, the user must create a custom, combinatorial
naming scheme to avoid conflicts. They typically require root
privileges and do not optimize for specific hardware.

@@

Packaging problems arise in HPC because a supercomputer's hardware, OS, and
filesystem are shared by many users with different requirements.  The classic
solution to this problem is to use virtual
machines (VMs)~\cite{barham2003xen,rosenblum1999vmware,smith2005architecture}
or lightweight virtualization techniques like Linux
containers~\cite{felter2014updated,merkel2014docker}. This model allows each
 user to have a personalized environment with its own package manager, and it
has been extremely successful for servers at cloud data centers. VMs typically
have near-native compute performance but low-level HPC network drivers still
exhibit major performance issues. VMs are not well supported on many
non-Linux operating systems, an issue for the lightweight
kernels of bleeding-edge Blue Gene/Q and Cray machines.
Finally, each VM still uses a traditional package manager,
so running many configurations still requires a large number of VMs.
For facilities, this is a security concern, as a profusion of VMs makes
mandatory patching difficult.  It is also tedious for users to manage a large
number of VM environments.

@@

To cope with software diversity, many HPC sites use a combination of existing
package managers and either manual or semi-automated installation.
For the baseline OS, many sites maintain traditional binary
packages using the vendor's package manager. LLNL maintains a Linux
distribution, CHAOS~\cite{chaos} for this purpose, which is managed using RPM.
%
For custom builds, many sites adhere to detailed naming conventions
that encode information in file system paths.
Table~\ref{tab:naming-conventions} shows several sites' conventions.
LLNL uses the APT package manager for installs
in the {\tt /usr/local/tools} file system and {\tt /usr/global/tools}
for manual installs.
ORNL uses hand installs but adheres to strict scripting conventions
to reproduce each build~\cite{jones+:cug08}.
TACC relies heavily on locally maintained RPMs.

@@

Recently, a number of HPC package managers have emerged that manage
multi-configuration builds.
%
Smithy~\cite{digirolamo:smithy} is an installation tool in use at ORNL. It
can generate module files, but it does not provide any
automated dependency management; it only checks whether a package's
prerequisites have already been installed by the user.

Nix~\cite{dolstra+:icfp08,dolstra+:lisa04}
is a package manager and an OS distribution that supports the installation of
arbitrarily many software configurations.  As at most HPC sites, each package
in Nix is installed in a unique prefix, but Nix does not have a human-readable
naming convention.  Instead, Nix determines the prefix by hashing the package
file along with its dependencies. Nix package files are written in a custom
functional language designed for packaging.

@@

\paragraph{Gaps in current practice}
The cryptographic versioning of Nix and HashDist is very flexible. It versions
the package {\it and} its dependency configuration, and can represent any
configuration. However, users cannot navigate or easily query the installed
software.
%The systems are in a sense ``write-only'':
%Nix does not offer a way to view a package's dependencies.
%
EasyBuild and Smithy generate environment modules, and some querying is
possible through the module system.  Naming schemes used in existing module
systems, however, cannot handle combinatorial versions. The Lmod authors
call this the ``matrix problem''~\cite{mclay:lmod-tutorial}.

@@

The main limitation of existing tools is the lack of build {\it composability}.
The full set of package versions is combinatorial, and realizing an arbitrary
combination of compiler, MPI version, build options, and dependency versions
requires tediously modifying many package files.
Indeed, the number of package files required for most existing systems scales
with the number of version {\it combinations}, not the number of packages, which
quickly becomes unmanageable.  As an example, the EasyBuild
system has over 3,300 files for several permutations of only 600 packages.
A slightly different dependency graph requires an entire new package file
hierarchy.  HashDist a more robust support for composition,
but it does not have a first-class parameters for versions, compilers,
or versioned interfaces.
%
HPC sites need better mechanisms to {\it parameterize} packages so that new
builds can be {\it composed} in response to user needs.
%

>>>>

\paragraph{Traditional package managers}
Package managers automate the installation of complex sets of software packages.
{\it Binary package managers} such as RPM, yum, and APT~\cite{foster+:rpm03,silva:apt01,yum} are integrated with most
OS distributions, and they are used to ensure that dependencies
are installed before packages that require them.
Others, such as Anaconda~\cite{anaconda,conda}, use the same approach but
are designed to run on top of a host OS.
These tools largely solve the problem of managing a {\it single} software
stack, which works well for the baseline OS and drivers, which are
common to all applications on a system.
These tools assume that each package only has a single version
and most of the tools install packages in a single, inflexible location.
To install multiple configurations, the user must create a custom, combinatorial
naming scheme to avoid conflicts. They typically require root
privileges and do not optimize for specific hardware.

@@

Packaging problems arise in HPC because a supercomputer's hardware, OS, and
filesystem are shared by many users with different requirements.  The classic
solution to this problem is to use virtual
machines (VMs)~\cite{barham2003xen,rosenblum1999vmware,smith2005architecture}
or lightweight virtualization techniques like Linux
containers~\cite{felter2014updated,merkel2014docker}. This model allows each
 user to have a personalized environment with its own package manager, and it
has been extremely successful for servers at cloud data centers. VMs typically
have near-native compute performance but low-level HPC network drivers still
exhibit major performance issues. VMs are not well supported on many
non-Linux operating systems, an issue for the lightweight
kernels of bleeding-edge Blue Gene/Q and Cray machines.
Finally, each VM still uses a traditional package manager,
so running many configurations still requires a large number of VMs.
For facilities, this profusion of VMs is a security concern since it 
complicates mandatory patching. Also, managing  a large number of VM 
environments is tedious for users.

@@

To cope with software diversity, many HPC sites use a combination of existing
package managers and either manual or semi-automated installation.
For the baseline OS, many sites maintain traditional binary
packages using the vendor's package manager. LLNL maintains a Linux
distribution, CHAOS~\cite{chaos} for this purpose, which is managed using RPM.
The popular ROCKS~\cite{rocks} cluster distribution uses RPM and Anaconda
in a similar fashion.
%
For custom builds, many sites adhere to detailed naming conventions
that encode information in file system paths.
Table~\ref{tab:naming-conventions} shows several sites' conventions.
LLNL uses the APT package manager for installs
in the {\tt /usr/local/tools} file system and {\tt /usr/global/tools}
for manual installs.
ORNL uses hand installs but adheres to strict scripting conventions
to reproduce each build~\cite{jones+:cug08}.
TACC relies heavily on locally maintained RPMs.

@@

Recently, a number of HPC package managers have emerged that manage
multi-configuration builds.
%
ORNL uses the Smithy~\cite{digirolamo:smithy} installation tool. It
can generate module files, but it does not provide any
automated dependency management; it only checks whether a package's
prerequisites have already been installed by the user.

The Nix~\cite{dolstra+:icfp08,dolstra+:lisa04} package manager and OS 
distribution supports installation of arbitrarily many software 
configurations.  As at most HPC sites, it installs each package
in a unique prefix but it does not have a human-readable naming 
convention.  Instead, Nix determines the prefix by hashing the package
file and its dependencies.
%% BRONIS: I'm not sure what the following sentence really adds
%%  Nix package files are written in a custom
%% functional language designed for packaging.

@@

\paragraph{Gaps in current practice}
The flexible cryptographic versioning of Nix and HashDist manages 
the package {\it and} its dependency configuration and can represent any
configuration. However, users cannot easily navigate or query the installed
software.
%The systems are in a sense ``write-only'':
%Nix does not offer a way to view a package's dependencies.
%
EasyBuild and Smithy generate environment modules, which supports some 
querying. Naming schemes used in existing module systems, however, cannot 
handle combinatorial versions, which the Lmod authors call the 
``matrix problem''~\cite{mclay:lmod-tutorial}.

@@

The main limitation of existing tools is the lack of build {\it composability}.
The full set of package versions is combinatorial, and arbitrary combinations 
of compiler, MPI version, build options, and dependency versions requires 
tedious modifications of many package files. Indeed, the number of package 
files required for most existing systems scales with the number of version 
{\it combinations}, not the number of packages, which quickly becomes 
unmanageable.  As an example, the EasyBuild system has over 3,300 files for 
several permutations of only 600 packages. A slightly different dependency 
graph requires an entire new package file hierarchy.  HashDist supports 
composition more robustly but does not have first-class parameters for 
versions, compilers, or versioned interfaces.
%
HPC sites need better mechanisms to {\it parameterize} packages so that new
builds can be {\it composed} in response to user needs.
%
